Local Tuning in Multilevel Scheme of Parallel Global Optimization

### 2016-05-02 2016/NUMTA/paper.tex
Finish tex version of NUMTA paper

% Document starts
\begin{document}

% Title portion
\title{Local Tuning in Multilevel Scheme of Parallel Global Optimization}

@@

Second, we shall assume that $\phi(y)$ satisfies Lipschitz condition
\begin{displaymath}
|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
that is typical for the applied problems [1–4].

Solving the multiextremal optimization problems is significantly more computationally intensive as compared with other types of optimization problems, since the global optimum is an integral characteristic of the problem solved and requires an analysis of the whole search domain. The efficient algorithms for solving the problems of this kind have been developed in Lobachevsky State University of Nizhnii Novgorod within the framework of the information-statistical approach [5, 6]. At present, the parallelization of the algorithms is the main direction of further development of these ones since the use of the supercomputer systems and of the parallel computing technologies allows expanding the field of application of the global optimization algorithms essentially [7 – 12].

This paper continues the series of studies the initial results of which were provided in [13,14].

@@

% Head 2
\subsection{DIMENSION REDUCTION}
There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in [15,16] or the simplicial partitions method in [17].

@@

In scheme (\ref{nestedScheme}), the nested subproblems are the multidimensional ones, and the dimension reduction method based on Peano curves can be applied for solving of these ones. The number of vectors $u_i$ and the quantity of components in each vector are the parameters of the block recursion scheme and can be utilized for the forming of the subproblems with necessary properties. For example, if $M=N$ i. e. $u_i=u_i,1\leqslant i\leqslant N$ , then each nested subproblem is a one-dimesional one. If $M=1$ i. e. $u=u_1=y$ , then solving the problem is an equivalent to its solving using a single evolvent mapping $[0,1]$ onto $D$; the nested subproblems are absent.
The issues of the numerical construction of the Peano curve approximations and of using the nested optimization scheme are considered in details in [3,4].

@@

A detailed description of various variants of this algorithm is presented in [5, 18, 19, 20].

@@

Various variants of the parallel algorithms based on this idea are presented in [7-12].

% Head 3
\subsubsection{MIXED LOCAL-GLOBAL ALGORITHM}
A scheme aimed at the acceleration of the search process is outlined in this subsection. The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end points of the corresponding interval. Following this idea, the “local” characteristics of the intervals can be introduced [5, 21].

@@

The problems generated by GKLS generator [22] were used as the test ones. This generator allows generating the multiextremal optimization problems with the properties known in advance: number of local minima, size of attraction domains, global minimum point, function values, etc. The experiment included solving a series of 200 problems with $N=4$ and $N=5$. The global minimum $y^*$ was considered to be found, if the algorithm generates a trial point $y^k$ in its $\delta$-vicinity $\Vert y^k-y^*\Vert\leqslant \delta$. The size of the vicinity was selected as $\delta=0.01\Vert b- a\Vert$ where $a$ and $b$ were the borders of the search domain $D$. The maximum allowable number of trials was $K_{max} = 10^6$.

>>>>

\begin{document}

\title{Local Tuning in Multilevel Scheme of Parallel Global Optimization}

@@

Second, we shall assume that $\phi(y)$ satisfies Lipschitz condition
\begin{displaymath}
|\phi(y_1)-\phi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
that is typical for the applied problems \citep{automaticaAppl,svmAppl,recogAppl,determApproaches}.

Solving the multiextremal optimization problems is significantly more computationally intensive as compared with other types of optimization problems, since the global optimum is an integral characteristic of the problem solved and requires an analysis of the whole search domain. The efficient algorithms for solving the problems of this kind have been developed in Lobachevsky State University of Nizhnii Novgorod within the framework of the information-statistical approach \citep{strongSerg,strongSergLera}. At present, the parallelization of the algorithms is the main direction of further development of these ones since the use of the supercomputer systems and of the parallel computing technologies allows expanding the field of application of the global optimization algorithms essentially \citep{parAlg,parAlg2,gergelLNCS,gergelStrParComputing,mixedAlgLNCS,barkXeon}.

This paper continues the series of studies the initial results of which were provided in \citep{barkGergNestedScheme,localTuning}.

@@

\subsection{DIMENSION REDUCTION}
There is a number of ways to adapt effective one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \citep{sergKvaDiagNew,kvasovDiag} or the simplicial partitions method in \citep{DISIMPL}.

@@

In scheme (\ref{nestedScheme}), the nested subproblems are the multidimensional ones, and the dimension reduction method based on Peano curves can be applied for solving of these ones. The number of vectors $u_i$ and the quantity of components in each vector are the parameters of the block recursion scheme and can be utilized for the forming of the subproblems with necessary properties. For example, if $M=N$ i. e. $u_i=u_i,1\leqslant i\leqslant N$ , then each nested subproblem is a one-dimesional one. If $M=1$ i. e. $u=u_1=y$ , then solving the problem is an equivalent to its solving using a single evolvent mapping $[0,1]$ onto $D$; the nested subproblems are absent.
The issues of the numerical construction of the Peano curve approximations and of using the nested optimization scheme are considered in details in \citep{recogAppl,determApproaches}.

@@

A detailed description of various variants of this algorithm is presented in \cite{strongSerg,gergelDerivMethod,branchBoundAlg,barkAdaptConstr}.

@@

Various variants of the parallel algorithms based on this idea are presented in \citep{parAlg,parAlg2,gergelLNCS,gergelStrParComputing,mixedAlgLNCS,barkXeon}.

% Head 3
\subsubsection{MIXED LOCAL-GLOBAL ALGORITHM}
A scheme aimed at the acceleration of the search process is outlined in this subsection. The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end points of the corresponding interval. Following this idea, the “local” characteristics of the intervals can be introduced \citep{strongSerg,indexMethod}.

@@

The problems generated by GKLS generator \citep{GKLS} were used as the test ones. This generator allows generating the multiextremal optimization problems with the properties known in advance: number of local minima, size of attraction domains, global minimum point, function values, etc. The experiment included solving a series of 200 problems with $N=4$ and $N=5$. The global minimum $y^*$ was considered to be found, if the algorithm generates a trial point $y^k$ in its $\delta$-vicinity $\Vert y^k-y^*\Vert\leqslant \delta$. The size of the vicinity was selected as $\delta=0.01\Vert b- a\Vert$ where $a$ and $b$ were the borders of the search domain $D$. The maximum allowable number of trials was $K_{max} = 10^6$.
